\section{DAG}

%%%%%%%%%%%%%%%
\begin{frame}{DAG}
  \begin{center}
    no back edge $\iff$ DAG $\iff$ $\exists$ topo. ordering
  \end{center}

  \begin{block}{Topo. sorting}
    DFS on digraph, $u \to v$:
    \begin{itemize}
      \item back edge: $f(u) < f(v)$
      \item others: $f(u) > f(v)$
    \end{itemize}

    \[ u \to v \Rightarrow u \prec v \] 
    \[ u \to v \Rightarrow f(u) > f(v) \]

    \begin{center}
      Topo. sorting: sort vertices in \textcolor{blue}{\emph{decreasing}} order of their \textcolor{blue}{\emph{finish}} times.
    \end{center}
  \end{block}
\end{frame}
%%%%%%%%%%%%%%%
\begin{frame}{Digraph as DAG}
  \begin{exampleblock}{Digraph as DAG \pno{3.4.6}}
    \begin{theorem}
      Every digraph is a dag of its SCCs.
    \end{theorem}
  \end{exampleblock}

  \begin{alertblock}{Remark.}
    \begin{itemize}
      \item SCC algorithm
      \item SCC: reachability/connectivity equivalence class
      \item two tiered structure of digraphs
    \end{itemize}
  \end{alertblock}
\end{frame}
%%%%%%%%%%%%%%%
\begin{frame}{Kahn's toposort algorithm}
  \begin{exampleblock}{Kahn's toposort algorithm (1962) \pno{3.4.19}}
    \begin{itemize}
      \item queue for source vertices ($\text{in}[v] = 0$)
      \item repeat: dequeue $v$, delete it, output it
    \end{itemize}
  \end{exampleblock}

  \begin{block}{Solution.}
    \begin{lemma}
      Every DAG has at least one source and at least one sink vertex.
    \end{lemma}
  \end{block}

  \begin{alertblock}{Remark}
    DFS on DAG:
    \begin{itemize}
      \item $\argmax_{v} f(v)$ $\Rightarrow$ source (used in SCC algorithm)
      \item $\argmin_{v} f(v)$ $\Rightarrow$ sink
    \end{itemize}
  \end{alertblock}
\end{frame}
%%%%%%%%%%%%%%%
\begin{frame}{Hamiltonian path in DAG}
  \begin{exampleblock}{Hamilton path in DAG \pno{3.4.16}}
    \begin{itemize}
      \item DAG $G$
      \item path visiting each vertex once
    \end{itemize}
  \end{exampleblock}

  \begin{block}{Solution.}
    \begin{itemize}
      \item general digraph: NP-hard
      \item dag: $\exists$ HP $\iff$ $\exists!$ topo. ordering
	\begin{itemize}
	  \item $\Leftarrow$: By contridiction. $\exists u \sim v: u \nrightarrow v$; swap
	\end{itemize}
    \end{itemize}

    Algorithm:
    \begin{itemize}
      \item toposort, check edges
      \item the Kahn toposort algorithm
    \end{itemize}
  \end{block}
\end{frame}
%%%%%%%%%%%%%%%
\begin{frame}{Semi-connected DAG}
  \begin{definition}{Semi-connected digraph}
    $\forall u,v: u \leadsto v \lor v \leadsto u$
  \end{definition}

  \begin{exampleblock}{Semi-connected DAG \pno{3.4.21 (c) + (d)}}
    To test whether a DAG is semi-connected.
  \end{exampleblock}

  \begin{block}{Solution.}
    \begin{center}
      dag: $\exists$ HP $\iff$ $\exists!$ topo. ordering $\iff$ semi-connected
    \end{center}

    \begin{proof}
      \begin{itemize}
	\item $\Leftarrow$: by contradiction; total order ($\forall u,v: u \prec v \lor v \prec u$)
	\item $\Rightarrow$: $\exists HP$
      \end{itemize}
    \end{proof}
  \end{block}
\end{frame}
%%%%%%%%%%%%%%%
\begin{frame}{Minimum cost reachable}
  \begin{exampleblock}{Minimun cost reachable \pno{3.4.22}}
    Compute $\text{cost}[u] = \min \set{\text{cost}[v] \mid u \leadsto v}$.
  \end{exampleblock}

  \begin{block}{Solution.}
    \begin{itemize}
      \item dag: reverse topo. ordering
	\begin{itemize}
	  \item backtracking: $\text{cost}[u] = \min_{u \to v} \set{\text{cost}[v]}$
	\end{itemize}
      \item digraph: dag of scc
    \end{itemize}
  \end{block}
\end{frame}
%%%%%%%%%%%%%%%
\begin{frame}{Line up}
  \begin{exampleblock}{Line up \pno{3.4.29}}
    \begin{itemize}
      \item $i$ hates $j$: $i \prec j$
      \item $i$ hates $j$: $\# i < \# j$
    \end{itemize}
  \end{exampleblock}

  \begin{block}{Solution.}
    $i$ hates $j$: $i \to j$;
    \begin{itemize}
      \item DAG?
      \item longest path; critical path
      \end{itemize}
  \end{block}
\end{frame}
%%%%%%%%%%%%%%%
\begin{frame}{One-to-all reachability}
  \begin{exampleblock}{One-to-all reachability \pno{3.4.28}}
    \begin{itemize}
      \item given $v: v \leadsto^{?} \forall u$
      \item $\exists? v: v \leadsto \forall u$
    \end{itemize}
  \end{exampleblock}

  \begin{block}{Solution.}
    \begin{itemize}
      \item DFS/BFS
      \item SCC; $\exists!$ source vertex $v$ $\iff$ $v \leadsto \forall u$
	\begin{proof}
	  \begin{itemize}
	    \item $\Rightarrow$: By contradiction. $\exists u: v \nrightarrow u \land \text{in}[u] > 0 \Rightarrow \exists u' \to u \land v \nrightarrow u'$. Cycle. 
	    \item $\Leftarrow$: (1) source; (2) $\exists !$
	  \end{itemize}
	\end{proof}
    \end{itemize}
  \end{block}
\end{frame}
%%%%%%%%%%%%%%%
